---
title: "Ulva and Hypnea Photosynthesis and GROWTH Run5 - Sept-Nov 2021"
author: "Angela Richards Donà"
date: "1/28/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Run5 GROWTH RATE Analysis, Script Chunks, and Plots

This is the analysis of the final run of the Ulva and Hypnea salinity and nutrient experiments 
conducted on the lanai in St. John 616. These experiments incorporated three temperature levels.

Packages loaded:

```{r, message=FALSE}
library(lme4)
library(lmerTest)
library(effects)
library(car)
library(MuMIn)
library (dplyr)
library(emmeans)
library(DHARMa)
library(performance)
library(patchwork)
```
## Load and prepare the dataset
Open growth/weight dataset

```{r}
run5_growth <- read.csv("/Users/Angela/Library/Mobile Documents/com~apple~CloudDocs/research_limu/run5_hyp_ulv_temp/final_temperature_run1-4/final_data_files/run5_growth_all_032422.csv")
```

Make a new column for weight change (difference final from initial)

```{r run5_growth}
run5_growth$growth_rate_percent <- 
  (run5_growth$final.weight - run5_growth$Inital.weight) / run5_growth$Inital.weight * 100
```

Also make a new column for daily growth rate from 8 day study (steady growth rate assumed rather than exponential), 
which may or may not be used

```{r}
run5_growth$steady_growth_daily <- run5_growth$growth_rate_percent / 8
```
## Make a new column that keeps only the numerical values for temperature (removes C)

```{r}
run5_growth$temp_clean <- as.factor(substr(run5_growth$temperature, 1, 2))
```
## Change levels to factors

```{r}
run5_growth$temperature <- as.factor(run5_growth$temp_clean)
run5_growth$run <- as.factor(run5_growth$run)
```

```{r}
run5_growth$treatment <- as.factor(as.character(run5_growth$treatment))

```
#assign run as a factor

```{r}
run5_growth$run <- as.factor(run5_growth$run)
```

# Create subset of the data to isolate the species
```{r}
hypnea <- subset(run5_growth, Species == "Hm")
ulva <- subset(run5_growth, Species == "Ul")
```

# Run the model

# ULVA 
# run model without interaction since interaction caused collinearity issues

```{r ulva}
run5_growth_model_noint <- lmer(formula = growth_rate_percent ~ treatment + temperature + (1 | run), 
                                data = ulva)
```

#make a histogram of the data for Ulva

```{r run5_growth_model_noint, fig.width=5}
hist(ulva$growth_rate_percent, main = paste("Ulva lactuca Growth Rate (%)"),
     col = "olivedrab3", labels = TRUE)
```
```{r}
plot(resid(run5_growth_model_noint) ~ fitted(run5_growth_model_noint))
qqnorm(resid(run5_growth_model_noint))
qqline(resid(run5_growth_model_noint))
```

## Check the performance of the model for Ulva
```{r fig.height=8}
performance::check_model(run5_growth_model_noint)
```
These outputs show the model is acceptable

##Run ANOVA and Tukey's comparison
```{r}
anova(run5_growth_model_noint, type = c("III"), ddf = "Satterthwaite")
ulva_growth_model_aov <- aov(growth_rate_percent ~ treatment + temperature, data = ulva)
TukeyHSD(ulva_growth_model_aov, "treatment", ordered = FALSE)
```

```{r}
r.squaredGLMM(run5_growth_model_noint)
summary(run5_growth_model_noint)
```

# Effects Plots
```{r}
plot(allEffects(run5_growth_model_noint))
```



# HYPNEA 
```{r hypnea}
run5_growth_model_noint <- lmer(formula = growth_rate_percent ~ treatment + temperature + (1 | run),
                                data = hypnea)
```

## Check the performance of the model for Hypnea
```{r, fig.height=8}
performance::check_model(run5_growth_model_noint)
```
These outputs show the model is acceptable for the data

```{r, fig.width=5}
hist(hypnea$growth_rate_percent, main = paste("Hypnea musciformis Growth Rate (%)"), 
     col = "maroon", labels = TRUE)
```
```{r}
plot(resid(run5_growth_model_noint) ~ fitted(run5_growth_model_noint))
qqnorm(resid(run5_growth_model_noint))
qqline(resid(run5_growth_model_noint))
```

```{r}
anova(run5_growth_model_noint, type = c("III"), ddf = "Satterthwaite")
hypnea_growth_model_aov <- aov(growth_rate_percent ~ treatment + temperature, data = hypnea)
TukeyHSD(hypnea_growth_model_aov, "treatment", ordered = FALSE)
```

```{r}
r.squaredGLMM(run5_growth_model_noint)
summary(run5_growth_model_noint)

```
# Effects Plots
```{r}
plot(allEffects(run5_growth_model_noint))


```
